### 2025-05-09 manuscript/manuscript.Rmd
Very rough first draft of the results

---
title: ""
author:
  - William Kumler
  - Laura T. Carlson
  - Anitra E. Ingalls
output: 
  word_document:
    reference_docx: Frontiers_Template.docx
bibliography:
- Exported Items.bib
- r-packages.bib
csl: frontiers-in-marine-science.csl
---

@@

```{r setup}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE)
library(tidyverse)
library(ggh4x)
library(patchwork)
library(ggtext)
library(minpack.lm)
options(pillar.sigfig = 7, pillar.print_max = 100)

parametadata <- read_csv("metadata/parametadata.csv") %>%
  # filter(IS_type=="noIS") %>%
  filter(polarity=="pos") %>%
  filter(samp_type=="Smp") %>%
  mutate(startime=factor(startime, levels=c("Morn", "Eve"))) %>%
  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
  mutate(amendment=factor(amendment, levels=c("Amm", "Nit", "Arg", "GMP"))) %>%
  arrange(amendment, depth, timepoint_num, startime) %>%
  mutate(timepoint=fct_inorder(timepoint))
peak_areas <- read_csv("targeted/all_peak_areas_w_isos.csv") %>%
  left_join(parametadata %>% distinct(filename, shortname))
all_stans <- read_csv("metadata/all_stans.csv")
maa_names <- c("Also Palythene/Usujirene?", "Shinorine?", "Mycosporine-glycine?", 
               "Palythine?", "Another Palythene/Usujirene?", "Porphyra-334?", 
               "Also Porphyra-334?", "Asterina-330?", "Palythene/Usujirene?",
               "Mycosporine-2-glycine?")
nucleo_names <- c("5-Methylcytosine", "Adenosine", "Cytidine", "Adenine", "Guanosine",
                  "Cytosine", "Hypoxanthine/Allopurinol", "Guanine", "Deoxycytidine?",
                  "Deoxyadenosine", "Deoxyguanosine?")
betaine_names <- c("beta-Alaninebetaine", "Glycine betaine", "Homarine", "Trigonelline",
                   "Choline", "Glycerophosphocholine", "Homoserine betaine?", 
                   "Threonine betaine?", "Proline betaine",
                   "(3-Carboxypropyl)trimethylammonium", "O-Acetylcarnitine", 
                   "Betonicine", "Acetylcholine", "Phosphocholine")
non_n_names <- c("Dimethylsulfonioacetate", "Dimethylsulfoniopropionate",
                 "Arsenobetaine", "Dimethylsulfoxonium propionate?", "Gonyol",
                 "2-O-alpha-D-Glucosylglycerol")
aa_names <- c("Glutamine", "Aspartic acid", "Glutamic acid/beta-Glutamic acid",
              "Asparagine", "Alanine", "Methionine", "Arginine", "Proline",
              "Lysine", "Threonine/Homoserine", "Serine", "Glycine", "Histidine",
              "Leucine/Isoleucine", "Valine", "Arginine")
bad_cmpds <- c("Melamine", "Uridine", "Itaconic acid?", "Inosine", "Dexpanthenol", 
               "Choline sulfate", "5-Hydroxyectoine", "L-Tyrosine")
samp_areas <- parametadata %>%
  filter(samp_type=="Smp") %>%
  distinct(shortname, depth) %>%
  left_join(peak_areas) %>%
  mutate(shortname=fct_inorder(shortname)) %>%
  filter(!compound_name%in%bad_cmpds) %>%
  filter(!(depth=="Deep" & compound_name%in%maa_names)) %>%
  select(-depth)
labeled_fracs <- samp_areas %>%
  mutate(labeled_frac=1-area/sum(area), .by = c("shortname", "compound_name")) %>%
  select(compound_name, iso_name, shortname, labeled_frac) %>%
  filter(str_detect(iso_name, "13C0, 15N0")) %>%
  select(-iso_name)
log_coef_df <- labeled_fracs %>%
  filter(!compound_name%in%maa_names) %>%
  left_join(parametadata %>% distinct(shortname, amendment, depth, timepoint_num)) %>%
  nest(data=c(shortname, timepoint_num, labeled_frac)) %>%
  mutate(log_fit=map(data, ~glm(.x$labeled_frac~.x$timepoint_num, family="quasibinomial")),
         coefs=map(log_fit, function(x)broom::tidy(x))) %>%
  unnest(coefs) %>%
  select(compound_name, amendment, depth, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  rename(intercept=`(Intercept)`, slope=`.x$timepoint_num`) %>%
  mutate(half_max=-intercept/slope, scale=1/slope) %>%
  mutate(frac_at_T24 = 1/(1+exp(-(intercept+slope*24))))

@@

## Abstract

## Introduction

Marine carbon is fixed into biomass at a rate equivalent to ten fully loaded Boeing 747 airplanes crashing into the ocean every second of every day [ref 50Pg IPCC(?), ref Boeing weights]. In most of the surface ocean this flux is controlled by the availability of nitrogen, an element with a complex and incompletely characterized biogeochemical cycle [@Capone2008; @Moore2013; @Hutchins2022]. Organic nitrogen in particular is often treated as a black box despite substantial variability in its bioavailability and chemical nature. While extensive work has been done to characterize the forms and fluxes of nitrogen within the inorganic pool, the equivalent for the organic pool is woefully underdeveloped [@ref, prob NME2008 or Hansell2024, maybe @Hutchins2022 for visual?]. This is largely due to the difficulty of comprehensively measuring the many organic molecules that contain nitrogen in the marine environment [@Boysen2018; @Moran2022]. The bioavailability of the nitrogen atom is in the environment is largely determined by the form of the organic material it composes.

@@

### Other things to maybe discuss:

@@

### Other soundbites:
  - It is assumed that most fixed nitrogen becomes protein, presumably based on the large fraction of cell nitrogen in protein and previous research showing how quickly it enters the (free?) amino acid pool. However, 1) important intermediates molecules 2) doesn’t explain where biounavailable N comes from 3) makes gene-based molecular modeling really difficult.
  - In each case, assimilation into organic matter is typically done via the glutamine synthetase (GS)/glutamate synthase (GOGAT) or glutamate dehydrogenase (GDH) pathways to produce glutamate from α-ketoglutarate [@ref, see bronk chapter]. Glutamate then serves as a nitrogen supply in an enormous number of biochemical pathways in both primary and secondary metabolism [@Walker2016].
  - nitrate amendments typically resulting in diatom blooms while ammonia incubations favor cyanobacteria [@Glibert2016].
  - Communities to distinguish are 1) surface (largely regenerated production) and 2) 175m (largely new production(??))
  - use of organics vs synthesis of organics
  - Compounds such as non-proteinogenic amino acids (e.g., ornithine, citrulline, creatine, MAAs), osmolytes like betaines, nucleobases, and sulfur-containing molecules like taurine represent significant but poorly quantified pools. Understanding the rates and pathways by which these compounds are utilized is crucial, as they form the fundamental unit linking genetic potential to elemental cycling in biogeochemical models.
  - Nitrogen use varies significantly between the sunlit surface where productivity is largely regenerated through recycling of ammonia and organic nitrogen, and deeper waters near the base of the euphotic zone, where nitrate supports “new” production. Quantifying rates of nitrogen assimilation into ammonia and downstream products across these gradients is therefore essential for predicting marine productivity and nutrient cycling.
  - Nonetheless, how the community uses a substrate is arguably more important than whether it can.

## Results

### Environmental context of PARAGON cruise

Make plots from Rhea's data? Would love to add her as collaborator

Maybe just ref to PARAGON synthesis paper, should include Angel's IFCB data showing mostly Hemiaulus and its disappearance over the course of the cruise

Ref to Fuyan's published paper or remake the graph here?

### Rapid uptake and use of all N substrates

```{r}
env_concs <- read_csv("targeted/final_env_concs.csv")
env_concs %>%
  mutate(mean_nm=mean(conc_nm_in_env), .by=compound_name) %>%
  arrange(desc(mean_nm)) %>%
  mutate(compound_name=ifelse(compound_name=="L-Glutamic acid/beta-Glutamic acid", "Glutamate", compound_name)) %>%
  mutate(compound_name=ifelse(compound_name%in%unique(compound_name)[1:19], compound_name, "Other")) %>%
  mutate(compound_name=fct_inorder(compound_name)) %>%
  summarise(conc_nm=sum(conc_nm_in_env), .by=c(compound_name, filename)) %>%
  left_join(parametadata) %>%
  ggplot(aes(x=tripl, y=conc_nm, fill=compound_name)) +
  geom_col(color="black", width=1) +
  facet_nested(depth+startime~amendment+timepoint, switch="x") +
  scale_y_continuous(expand=expansion(mult=c(0, 0.1))) +
  scale_fill_manual(values = c(splividis(19), "grey")) +
  labs(x=NULL, y="Unlabeled concentrations (nM)", fill=NULL) +
  theme_bw() +
  theme(
    panel.spacing.x = unit(0, "lines"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
```

```{r}
peak_areas %>%
  filter(!str_detect(iso_name, "13C0, 15N0")) %>%
  filter(!compound_name%in%c("Urea", "L-Valine", "Oxalic acid", "Malic acid", 
                             "Acetylcholine", "Homarine")) %>%
  left_join(parametadata %>% distinct(filename, samp_type, amendment)) %>%
  filter(samp_type=="Smp") %>%
  left_join(read_csv("targeted/rf_df.csv")) %>%
  mutate(conc_nm=area/rf) %>%
  filter(!is.na(conc_nm)) %>%
  mutate(mean_nm=mean(conc_nm), .by=compound_name) %>%
  arrange(desc(mean_nm)) %>%
  mutate(compound_name=ifelse(compound_name=="L-Glutamic acid/beta-Glutamic acid", 
                              "Glutamate", compound_name)) %>%
  mutate(compound_name=ifelse(compound_name%in%unique(compound_name)[1:19], 
                              compound_name, "Other")) %>%
  mutate(compound_name=fct_inorder(compound_name)) %>%
  summarise(conc_nm=sum(conc_nm), .by=c(compound_name, filename)) %>%
  left_join(parametadata) %>%
  ggplot(aes(x=tripl, y=conc_nm, fill=compound_name)) +
  geom_col(color="black", width=1) +
  facet_nested(depth+startime~amendment+timepoint, switch="x") +
  scale_y_continuous(expand=expansion(mult=c(0, 0.1))) +
  scale_fill_manual(values = c(splividis(19), "grey")) +
  labs(x=NULL, y="Unlabeled concentrations (nM)", fill=NULL) +
  theme_bw() +
  theme(
    panel.spacing.x = unit(0, "lines"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
```

```{r}
cbind(
  peak_areas %>%
    # filter(iso_name=="L-Glutamine, 13C0, 15N0") %>%
    filter(iso_name=="Glycine betaine, 13C0, 15N0") %>%
    filter(str_detect(filename, "T0")) %>%
    select(filename, area),
  read_csv("targeted/wIS_areas.csv") %>%
    # filter(compound_name=="L-Glutamine") %>%
    filter(compound_name=="Glycine betaine") %>%
    filter(str_detect(filename, "T0")) %>%
    select(wIS_filename=filename, wIS_area=area)
) %>%
  # with(lm(wIS_area~area)) %>% summary()
  ggplot() +
  geom_point(aes(x=area, y=wIS_area)) +
  geom_abline(slope = 1.5, intercept = 0)
```

Total fraction of PN that was visible in metabolites, maybe big stacked barplot like usual?

Total number of compounds labeled over time (line plot?)

```{r thaa data}
thaa_quant <- read_csv("thaas/quant_data.csv")
metathaa <- read_csv("thaas/metathaa.csv") %>%
  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
  mutate(startime=factor(startime, levels=c("Morn", "Eve")))
aa_ncs <- all_stans %>%
  filter(compound_type=="Amino Acid") %>%
  mutate(compound_name=str_remove(compound_name, "^L-")) %>%
  distinct(compound_name, n_C, n_N) %>%
  add_row(compound_name="Phenylalanine", n_C=9, n_N=1)

![](figures/total_thaa_gp.png)
  - THAAs were 30-50% of total PC (~2 uM, need actual data) and 70-100% PN (~0.35uM)
  - Maybe more labeling happening from inorganics in evening?
  - 50% turnover of arginine in proteins in 3 days per +Arg data, less at surface than deep (??)
  - 30-50% Glu/Gln turnover in proteins in one day per Amm data
  - Glutamate, proline fully labeled labeled from arginine
  - Phenylalanine, tyrosine, histidine(?) labeled from GMP
  - How is there so much nitrate labeling when I don't see that in the metabolites?

### Universal bioavailability of ammonia at surface

@@

Synthesis rates?

@@

  - Carbon backbone traced into glutamate from arginine
  - Carbon backbone traced into other nucleobases from GMP
  - Possibly due to way too much of a given compound, useless to have that much Arg/GMP
  - Not very much fully-labeled guanine despite guanosine being easily cleavable, instead it’s mostly made from scratch (CHECK THIS)
  - Creatine for being mostly 2x15N 1x13C labeled from arginine

@@

Untargeted metabolites? Clustering their % labeled patterns with the targeted metabolites to see what groups they might belong to

## Discussion

## Conclusion

## Methods

## Acknowledgements

## Data availability

## References

```{r packages, eval=FALSE}
packages_used <- list.files("..", pattern = "*\\.R$|*\\.Rmd$", recursive = TRUE) %>%
  str_subset("old|scripts", negate=TRUE) %>%
  map(function(file_name){
    v <- readLines(paste0("../", file_name))
    libs <- unique(str_extract(v, "(?<=library\\().*?(?=\\)|,)"))
    refs <- unique(str_extract(v, "[[:alnum:]]*(?=::)"))
    pkgs <- unique(c(na.omit(libs), na.omit(refs)))
  }) %>% unlist() %>% unique()
knitr::write_bib(packages_used[nchar(packages_used)>0], file = "r-packages.bib")
```

@@

## Supplement

### Supp fig 1: full heatmap of all compounds and timepoints

@@

>>>>

---
title: ""
author:
  - William Kumler
  - Laura T. Carlson
  - Anitra E. Ingalls
output: 
  word_document:
    reference_docx: Frontiers_Template.docx
csl: frontiers-in-marine-science.csl
bibliography:
  - r-packages.bib
---

@@

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE)
library(tidyverse)
library(ggh4x)
library(patchwork)
library(ggtext)
library(minpack.lm)
options(pillar.sigfig = 7)
options(pillar.print_min = 100)

parametadata <- read_csv("metadata/parametadata.csv") %>%
  # filter(IS_type=="noIS") %>%
  filter(polarity=="pos") %>%
  filter(samp_type=="Smp") %>%
  mutate(startime=factor(startime, levels=c("Morn", "Eve"))) %>%
  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
  mutate(amendment=factor(amendment, levels=c("Amm", "Nit", "Arg", "GMP"))) %>%
  arrange(amendment, depth, timepoint_num, startime) %>%
  mutate(timepoint=fct_inorder(timepoint))
peak_areas <- read_csv("targeted/all_peak_areas_w_isos.csv") %>%
  left_join(parametadata %>% distinct(filename, shortname))
all_stans <- read_csv("metadata/all_stans.csv")
stan_patched <- all_stans %>%
  filter(!compound_name=="beta-Glutamic acid") %>%
  mutate(compound_name=str_replace(compound_name, "L-Glutamic acid", "L-Glutamic acid/beta-Glutamic acid")) %>%
  filter(!compound_name=="L-Isoleucine") %>%
  mutate(compound_name=str_replace(compound_name, "L-Leucine", "L-Leucine/L-Isoleucine")) %>%
  filter(!compound_name=="L-Homoserine") %>%
  mutate(compound_name=str_replace(compound_name, "L-Threonine", "L-Threonine/L-Homoserine")) %>%
  filter(!compound_name=="Allopurinol") %>%
  mutate(compound_name=str_replace(compound_name, "Hypoxanthine", "Hypoxanthine/Allopurinol"))

maa_names <- c("Also Palythene/Usujirene?", "Shinorine?", "Mycosporine-glycine?", 
               "Palythine?", "Another Palythene/Usujirene?", "Porphyra-334?", 
               "Also Porphyra-334?", "Asterina-330?", "Palythene/Usujirene?",
               "Mycosporine-2-glycine?")
nucleo_names <- c("5-Methylcytosine", "Adenosine", "Cytidine", "Adenine", "Guanosine",
                  "Cytosine", "Hypoxanthine/Allopurinol", "Guanine", "Deoxycytidine?",
                  "Deoxyadenosine", "Deoxyguanosine?")
betaine_names <- c("beta-Alaninebetaine", "Glycine betaine", "Homarine", "Trigonelline",
                   "Choline", "Glycerophosphocholine", "Homoserine betaine?", 
                   "Threonine betaine?", "Proline betaine",
                   "(3-Carboxypropyl)trimethylammonium", "O-Acetylcarnitine", 
                   "Betonicine", "Acetylcholine", "Phosphocholine")
non_n_names <- c("Dimethylsulfonioacetate", "Dimethylsulfoniopropionate",
                 "Arsenobetaine", "Dimethylsulfoxonium propionate?", "Gonyol",
                 "2-O-alpha-D-Glucosylglycerol")
aa_names <- c("Glutamine", "Aspartic acid", "Glutamic acid/beta-Glutamic acid",
              "Asparagine", "Alanine", "Methionine", "Arginine", "Proline",
              "Lysine", "Threonine/Homoserine", "Serine", "Glycine", "Histidine",
              "Leucine/Isoleucine", "Valine", "Arginine")
bad_cmpds <- c("Melamine", "Uridine", "Itaconic acid?", "Inosine", "Dexpanthenol", 
               "Choline sulfate", "5-Hydroxyectoine", "L-Tyrosine")
aa_ncs <- all_stans %>%
  filter(compound_type=="Amino Acid") %>%
  mutate(compound_name=str_remove(compound_name, "^L-")) %>%
  distinct(compound_name, n_C, n_N) %>%
  add_row(compound_name="Phenylalanine", n_C=9, n_N=1)

samp_areas <- parametadata %>%
  filter(samp_type=="Smp") %>%
  distinct(shortname, depth) %>%
  left_join(peak_areas) %>%
  mutate(shortname=fct_inorder(shortname)) %>%
  filter(!compound_name%in%bad_cmpds) %>%
  filter(!(depth=="Deep" & compound_name%in%maa_names)) %>%
  select(-depth)
labeled_fracs <- samp_areas %>%
  mutate(labeled_frac=1-area/sum(area), .by = c("shortname", "compound_name")) %>%
  select(compound_name, iso_name, shortname, labeled_frac) %>%
  filter(str_detect(iso_name, "13C0, 15N0")) %>%
  select(-iso_name)
log_coef_df <- labeled_fracs %>%
  filter(!compound_name%in%maa_names) %>%
  left_join(parametadata %>% distinct(shortname, amendment, depth, timepoint_num)) %>%
  nest(data=c(shortname, timepoint_num, labeled_frac)) %>%
  mutate(log_fit=map(data, ~glm(.x$labeled_frac~.x$timepoint_num, family="quasibinomial")),
         coefs=map(log_fit, function(x)broom::tidy(x))) %>%
  unnest(coefs) %>%
  select(compound_name, amendment, depth, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  rename(intercept=`(Intercept)`, slope=`.x$timepoint_num`) %>%
  mutate(half_max=-intercept/slope, scale=1/slope) %>%
  mutate(frac_at_T24 = 1/(1+exp(-(intercept+slope*24))))

@@

# Abstract

Nitrogen availability controls productivity in much of the surface ocean. The form of this nitrogen controls who it is available to and how it is used. Here, we explored how four different forms of nitrogen (ammonia, nitrate, arginine, and guanosine monophosphate) were taken up and transformed in the North Pacific Subtropical Gyre using stable isotope labeling metabolomics. We repeated these experiments both at the surface and at the base of the euphotic zone as well as in the morning and evening. We found that ammonia is converted into glutamate in less than 20 minutes and results in labeling throughout the metabolome, though osmolytes like betaines had slow turnover rates. This process was slower at depth but independent of time of day. Nitrate amendments resulted in the labeling of only a small subset of molecules such as mycosporine-like amino acids, whose incorporation patterns suggest that this form of nitrogen was available to a subset of the community, likely eukaryotes. The two organic nitrogen amendments labeled the metabolome more slowly than the inorganic forms of nitrogen and were primarily converted into glutamate rather than used as intact substrates. They also showed the same usage at depth as they did at the surface, hinting at the major role that heterotrophic bacteria likely play in their uptake and use. These measurements reveal the forms and fluxes of organic nitrogen in the natural environment to provide constraints on biogeochemical models and better characterize the transformations between organic and inorganic nitrogen.

# Introduction

Marine carbon is fixed into biomass very quickly. In most of the surface ocean this flux is controlled by the availability of nitrogen, an element with a complex and incompletely characterized biogeochemical cycle [@Capone2008; @Moore2013; @Hutchins2022]. Organic nitrogen in particular is often treated as a black box despite substantial variability in its bioavailability and chemical nature. While extensive work has been done to characterize the forms and fluxes of nitrogen within the inorganic pool, the equivalent for the organic pool is woefully underdeveloped [@ref, prob NME2008 or Hansell2024, maybe @Hutchins2022 for visual?]. This is largely due to the difficulty of comprehensively measuring the many organic molecules that contain nitrogen in the marine environment [@Boysen2018; @Moran2022]. The bioavailability of the nitrogen atom is in the environment is largely determined by the form of the organic material it composes.

@@

## Other things to maybe discuss:

@@

## Other soundbites:
  - It is assumed that most fixed nitrogen becomes protein, presumably based on the large fraction of cell nitrogen in protein and previous research showing how quickly it enters the (free?) amino acid pool. However, 1) important intermediates molecules 2) doesn’t explain where biounavailable N comes from 3) makes gene-based molecular modeling really difficult.
  - In each case, assimilation into organic matter is typically done via the glutamine synthetase (GS)/glutamate synthase (GOGAT) or glutamate dehydrogenase (GDH) pathways to produce glutamate from α-ketoglutarate [@ref, see bronk chapter]. Glutamate then serves as a nitrogen supply in an enormous number of biochemical pathways in both primary and secondary metabolism [@Walker2016].
  - nitrate amendments typically resulting in diatom blooms while ammonia incubations favor cyanobacteria [@Glibert2016].
  - Communities to distinguish are 1) surface (largely regenerated production) and 2) 175m (largely new production(??))
  - use of organics vs synthesis of organics
  - Compounds such as non-proteinogenic amino acids (e.g., ornithine, citrulline, creatine, MAAs), osmolytes like betaines, nucleobases, and sulfur-containing molecules like taurine represent significant but poorly quantified pools. Understanding the rates and pathways by which these compounds are utilized is crucial, as they form the fundamental unit linking genetic potential to elemental cycling in biogeochemical models.
  - Nitrogen use varies significantly between the sunlit surface where productivity is largely regenerated through recycling of ammonia and organic nitrogen, and deeper waters near the base of the euphotic zone, where nitrate supports “new” production. Quantifying rates of nitrogen assimilation into ammonia and downstream products across these gradients is therefore essential for predicting marine productivity and nutrient cycling.
  - Nonetheless, how the community uses a substrate is arguably more important than whether it can.

# Results

## Environmental context of PARAGON cruise

PARAGON synthesis paper isn't out yet and I'm not sure it will be anytime soon

Particulate carbon and nitrogen numbers for PARAGON I (from Matt(?))

![](figures/borrowed/pcpn_PIandPII.png)

Data from Rhea showing ammonia/nitrite concentrations - I sampled at 25m and 175m (nitrite max, below ammonia max). Does this mean I was getting a lot of AOA users?

![](figures/borrowed/rhea_ammonia_concs.png)

Angel's IFCB data showing Hemiaulus disappearing over the course of the cruise (most of the Bacillariophyceae are Hemiaulus and she shows that in another plot)

![](figures/borrowed/angel_hemiaulus.png)

Fuyan reports the abundances of various bacterial groups (with the caveat that the SASVs have to have been found in the 4000m depth traps?)

![](figures/borrowed/fuyan_composition_bact.png)

## Universal bioavailability of ammonia at surface

@@

Synthesis rates, especially for betaines?

@@

  - Carbon backbone traced into glutamate from arginine
  - Carbon backbone traced into other nucleobases from GMP
  - Possibly due to way too much of a given compound, useless to have that much Arg/GMP
  - Not very much fully-labeled guanine despite guanosine being easily cleavable, instead it’s mostly made from scratch (CHECK THIS)
  - Creatine for being mostly 2x15N 1x13C labeled from arginine (CHECK THIS)

@@

## Rapid uptake and use of all N substrates

Total fraction of PN that was visible in metabolites

```{r}
env_concs <- read_csv("targeted/final_env_concs.csv")

# More compounds, triplicates shown
env_concs %>%
  mutate(mean_nm=mean(conc_nm_in_env), .by=compound_name) %>%
  filter(!compound_name%in%c("Thymine", "Thymidine")) %>%
  arrange(desc(mean_nm)) %>%
  mutate(compound_name=ifelse(compound_name=="L-Glutamic acid/beta-Glutamic acid", "Glutamate", compound_name)) %>%
  mutate(compound_name=ifelse(compound_name%in%unique(compound_name)[1:19], compound_name, "Other")) %>%
  mutate(compound_name=fct_inorder(compound_name)) %>%
  summarise(conc_nm=sum(conc_nm_in_env), .by=c(compound_name, filename)) %>%
  left_join(parametadata) %>%
  ggplot(aes(x=tripl, y=conc_nm, fill=compound_name)) +
  geom_col(color="black", width=1) +
  facet_nested(depth+startime~amendment+timepoint, switch="x") +
  scale_y_continuous(expand=expansion(mult=c(0, 0.1))) +
  scale_fill_manual(values = c(splividis(19), "grey")) +
  labs(x=NULL, y="Unlabeled concentrations (nM)", fill=NULL) +
  theme_bw() +
  theme(
    panel.spacing.x = unit(0, "lines"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )

# Summarized version
env_concs %>%
  mutate(mean_nm=mean(conc_nm_in_env), .by=compound_name) %>%
  filter(!compound_name%in%c("Thymine", "Thymidine")) %>%
  arrange(desc(mean_nm)) %>%
  mutate(compound_name=ifelse(compound_name=="L-Glutamic acid/beta-Glutamic acid", 
                              "Glutamate", compound_name)) %>%
  mutate(compound_name=str_remove_all(compound_name, "^L-|(?<=/)L-")) %>%
  mutate(compound_name=ifelse(compound_name%in%unique(compound_name)[1:9], 
                              compound_name, "Other")) %>%
  mutate(compound_name=fct_inorder(compound_name)) %>%
  summarise(conc_nm=sum(conc_nm_in_env), .by=c(compound_name, filename)) %>%
  left_join(parametadata) %>%
  summarise(conc_nm=mean(conc_nm), .by=c(compound_name, amendment, depth, timepoint)) %>%
  ggplot(aes(x=timepoint, y=conc_nm, fill=compound_name)) +
  geom_col(color="black", width=1) +
  facet_grid(depth~amendment, scales="free") +
  scale_y_continuous(expand=expansion(mult=c(0, 0.1))) +
  scale_fill_manual(values = c(splividis(9), "grey")) +
  labs(x=NULL, y="Unlabeled concentrations (nM)", fill=NULL) +
  theme_bw()

# In N moles
env_concs %>%
  left_join(stan_patched %>% distinct(compound_name, n_N)) %>%
  mutate(nM_N=conc_nm_in_env*n_N) %>%
  mutate(mean_nm=mean(nM_N), .by=compound_name) %>%
  filter(!compound_name%in%c("Thymine", "Thymidine")) %>%
  arrange(desc(mean_nm)) %>%
  mutate(compound_name=ifelse(compound_name=="L-Glutamic acid/beta-Glutamic acid", 
                              "Glutamate", compound_name)) %>%
  mutate(compound_name=str_remove_all(compound_name, "^L-|(?<=/)L-")) %>%
  mutate(compound_name=ifelse(compound_name%in%unique(compound_name)[1:9], 
                              compound_name, "Other")) %>%
  mutate(compound_name=fct_inorder(compound_name)) %>%
  summarise(conc_nm=sum(nM_N), .by=c(compound_name, filename)) %>%
  left_join(parametadata) %>%
  summarise(conc_nm=mean(conc_nm), .by=c(compound_name, amendment, depth, timepoint)) %>%
  ggplot(aes(x=timepoint, y=conc_nm, fill=compound_name)) +
  geom_col(color="black", width=1) +
  facet_grid(depth~amendment, scales="free") +
  scale_y_continuous(expand=expansion(mult=c(0, 0.1))) +
  scale_fill_manual(values = c(splividis(9), "grey")) +
  labs(x=NULL, y="Nitrogen in unlabeled metabolites (nM)", fill=NULL) +
  theme_bw()
```

```{r}
peak_areas %>%
  filter(str_detect(iso_name, "13C0, 15N0")) %>%
  filter(!compound_name%in%c("Urea", "L-Valine", "Oxalic acid", "Malic acid", 
                             "Acetylcholine", "Homarine")) %>%
  left_join(parametadata %>% distinct(filename, samp_type, amendment)) %>%
  filter(samp_type=="Smp") %>%
  left_join(read_csv("targeted/rf_df.csv")) %>%
  mutate(conc_nm=area/rf) %>%
  filter(!is.na(conc_nm)) %>%
  mutate(mean_nm=mean(conc_nm), .by=compound_name) %>%
  arrange(desc(mean_nm)) %>%
  mutate(compound_name=ifelse(compound_name=="L-Glutamic acid/beta-Glutamic acid", 
                              "Glutamate", compound_name)) %>%
  mutate(compound_name=ifelse(compound_name%in%unique(compound_name)[1:19], 
                              compound_name, "Other")) %>%
  mutate(compound_name=fct_inorder(compound_name)) %>%
  summarise(conc_nm=sum(conc_nm), .by=c(compound_name, filename)) %>%
  left_join(parametadata) %>%
  ggplot(aes(x=tripl, y=conc_nm, fill=compound_name)) +
  geom_col(color="black", width=1) +
  facet_nested(depth+startime~amendment+timepoint, switch="x") +
  scale_y_continuous(expand=expansion(mult=c(0, 0.1))) +
  scale_fill_manual(values = c(splividis(19), "grey")) +
  labs(x=NULL, y="Unlabeled concentrations (nM)", fill=NULL) +
  theme_bw() +
  theme(
    panel.spacing.x = unit(0, "lines"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
```

## Labeling traced into protein via total hydrolyzable amino acids

```{r thaa data}
thaa_quant <- read_csv("thaas/quant_data.csv")
metathaa <- read_csv("thaas/metathaa.csv") %>%
  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
  mutate(startime=factor(startime, levels=c("Morn", "Eve")))

![](figures/total_thaa_gp.png)

  - THAAs were 30-50% of total PC (~2 uM, need actual data) and 70-100% PN (~0.35uM)
  - Maybe more labeling happening from inorganics in evening?
  - 50% turnover of arginine in proteins in 3 days per +Arg data, less at surface than deep (??)
  - 30-50% Glu/Gln turnover in proteins in one day per Amm data
  - Glutamate, proline fully labeled labeled from arginine
  - Phenylalanine, tyrosine, histidine(?) labeled from GMP
  - How is there so much nitrate labeling when I don't see that in the metabolites?

# Discussion

# Conclusion

# Methods

# Acknowledgements

# Data availability

# References

```{r packages, eval=FALSE}
packages_used <- list.files(pattern = "*\\.R$|*\\.Rmd$", recursive = TRUE, full.names = TRUE) %>%
  str_subset("old|scripts", negate=TRUE) %>%
  map(function(file_name){
    v <- readLines(file_name)
    libs <- unique(str_extract(v, "(?<=library\\().*?(?=\\)|,)"))
    refs <- unique(str_extract(v, "[[:alnum:]]*(?=::)"))
    pkgs <- unique(c(na.omit(libs), na.omit(refs)))
  }, .progress = TRUE) %>% unlist() %>% unique()
knitr::write_bib(packages_used[nchar(packages_used)>0], file = "manuscript/r-packages.bib")
```

@@

# Supplement

## Supp fig: full heatmap of all compounds and timepoints

@@

## Supp fig: THAA labeling broken down by triplicate
